---
title: "CorrelationPlot"
output: github_document
date: "2026-07-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function summary 

Calcultes the correlation coefficient between two variables (numeric or logical) 
on an LSOA level and plots on a map. Plotting donewith library sf and ggplot2 and
colour pallette from  grDevices::colorRampPalette. 

Current testing done on synthetic health outcome data created by RVD. 

```{r, include=FALSE, echo=FALSE}
    library(DSLite)
    library(devtools)
    library(dsBase)
    library(dsBaseClient)
    library(dsGeospatial)
    library(dsGeospatialClient)

# retrieve LSOA boundaries for Cheshire and Merseyside
# to reduce coverage, remove names from the `within_names`

cm_lsoa_sf <- boundr::bounds(
  "lsoa",
  within_level = "lad",
  within_names = c(
    "Cheshire East",
    "Cheshire West and Chester",
    "Halton",
    "Knowsley",
    "Liverpool",
    "Sefton",
    "St. Helens",
    "Warrington",
    "Wirral"
  ),
  lookup_year = 2011, # can be either 2011 or 2021
  opts = boundr::boundr_options(resolution = "BFC")
)

liverpool_sf <- cm_lsoa_sf |>
  dplyr::filter(lad22nm %in% "Liverpool")

uprn_gs_sf <- "https://pldr.org/download/2k6r3/n12/UPRN_2_1_greenspace_distances_with_coords.csv" |>
  readr::read_csv() |>
  # convert to spatial object
  sf::st_as_sf(coords = c("longitude", "latitude"), crs = 4326) |>
  # add LSOA details
  sf::st_join(cm_lsoa_sf)



qof_asthma_tbl <- "https://pldr.org/download/e6nzv/ng1/QOF_4_03_Asthma_LSOA.csv" |>
  readr::read_csv() |>
  # filter out CM LSOAs
  dplyr::filter(lsoa11 %in% cm_lsoa_sf$lsoa11cd) |>
  # filter out last year
  dplyr::filter(year == 2024)



# set seed for reproducibility
set.seed(20250603)
qof_asthma_cohort_tbl <- qof_asthma_tbl |>
  purrr::pmap(function(lsoa11, den, num, ...) {
    # recalculate pop to be a multiple of 2
    pop <- ifelse(den %% 2 != 0, den + 1, den)
    # extract UPRNs for this LSOA
    uprn_tbl <- uprn_gs_sf |>
      sf::st_drop_geometry() |>
      dplyr::filter(lsoa11cd == lsoa11) |>
      dplyr::mutate(
        dist_gs_dec = dplyr::ntile(distance_local_greenspace, 10)
      ) |>
      dplyr::arrange(dist_gs_dec)
    
    # create cohort table
    cohort_tbl <- tibble::tibble(
      lsoa11cd = lsoa11,
      sex = rep(c(0, 1), each = pop / 2),
      has_asthma = FALSE,
      uprn = sample(uprn_tbl$UPRN, size = pop, replace = TRUE)
    )
    
    # randomly assigned the outcome to `num` patients
    idx <- sample(seq_len(pop), ceiling(num), replace = FALSE)
    cohort_tbl$has_asthma[idx] <- TRUE
    
    # randomly re-assigned UPRNs with higher distances to patients with outcome
    ## create vector of probabilities, so further distances are favoured
    p <- c(0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.1, 0.2, 0.2, 0.2)
    idx_uprn_tile <- sample(1:10, ceiling(num), replace = TRUE, prob = p)
    ## subset UPRNs with match tile as per `idx_uprn_tile`
    cohort_tbl$uprn[idx] <- purrr::map(idx_uprn_tile, function(x) {
      # subset based on tile decile  
      aux <- dplyr::filter(uprn_tbl, dist_gs_dec == x)
      # extract UPRN
      aux$UPRN[sample(seq_len(nrow(aux)), 1)]
    }) |>
      purrr::list_c()
    
    # return final cohort table
    return(cohort_tbl)
  }) |>
  purrr::list_c()


qof_asthma_cohort_dist_gs_sf <- qof_asthma_cohort_tbl |>
  dplyr::left_join(uprn_gs_sf, by = c("uprn" = "UPRN", "lsoa11cd")) |>
  dplyr::arrange(lsoa11cd, uprn) |>
  sf::st_as_sf()
```


# Upload data data and load back with DSLite  

```{r , include=FALSE, echo=FALSE}
    
    options(nfilter.tab = 3)

    dat <- qof_asthma_cohort_dist_gs_sf |> 
      dplyr::select("lsoa11cd", "has_asthma","uprn","distance_local_greenspace") |> 
      sf::st_drop_geometry()

    ## ℹ Loading dsExampleClient

    dslite.server1 <- newDSLiteServer(
      tables = list(
        uprn_asthma_green1 = dat
      )
    )

    dslite.server2 <- newDSLiteServer(
      tables = list(
        uprn_asthma_green2 = dat
      )
    )


    dslite.server1$config(defaultDSConfiguration(include=c("dsBase", "dsExample")))
    dslite.server1$aggregateMethod("groupcorDS", "groupcorDS")

    dslite.server2$config(defaultDSConfiguration(include=c("dsBase", "dsExample")))
    dslite.server2$aggregateMethod("groupcorDS", "groupcorDS")

    builder <- DSI::newDSLoginBuilder()

    builder$append(
      server = "server_1",
      url = "dslite.server1",
      table = "uprn_asthma_green1",
      driver = "DSLiteDriver"
    )

    builder$append(
      server = "server_2",
      url = "dslite.server2",
      table = "uprn_asthma_green2",
      driver = "DSLiteDriver"
    )


    logindata <- builder$build()

    conns <- DSI::datashield.login(logins = logindata, assign = TRUE)

```

Set nfilter.tab 

```{r, eval=FALSE}
nfilter.tab  <- 3
```

## Output 
LSOA-level plotting can be done by combining or by splitting across the servers. 

```{r}
ds.groupcor("D$has_asthma", "D$distance_local_greenspace", "D$lsoa11cd", datasources = conns, type = "combine")
ds.groupcor("D$has_asthma", "D$distance_local_greenspace", "D$lsoa11cd", datasources = conns, type = "split")
```



