---
title: "Rukun's work"
output: html_document
date: "2026-08-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Study population characteristics

Before modelling the association between greenspace and asthma, we first describe the participating populations.

#### What does the distribution of variables look like?
--> Heatmap
We can first examine lsoa level summary statistics :
1. mean
2. Std deviation
3. Max
4. Min

Stats can be split/combined across study option to return plot/table
```{r, eval=FALSE}
ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns)
ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type='split')
ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type='split', names_region = c("Liverpool", "Sefton"))
ds.geoheatmapPlot(x="D$distance_local_greenspace", y = "D$lsoa11cd",
datasources = conns)
ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type = 'split', names_region = c("Liverpool", "Sefton"))
```
---> Histogram : may not be useful for lsoa level analysis

#### Is greenspace distance approximately normally distributed?
This is to be  developed
```{r, eval=FALSE}
ds.qqnorm()
```

#### Does asthma prevalence differ by sex/age/ethnicity/IMD?

For two categorical variables, we can use a chi-squared test.

```{r, eval=FALSE}
ds.chisqr(rvar="D$sex", cvar="D$has_asthma", stvar="D$lsoa11cd", 
          exclude=NULL,	useNA ="no", suppress.chisq.warnings=FALSE,
          datasources=NULL, names_region = c("Liverpool",  "Sefton"),
          force.nfilter=NULL, split = FALSE, draw.plot = TRUE)

ds.chisqr(rvar="D$sex", cvar="D$has_asthma", stvar="D$lsoa11cd", 
          exclude=NULL,	useNA ="no", suppress.chisq.warnings=FALSE,
          datasources=NULL, names_region = "Liverpool",
          force.nfilter=NULL, split = TRUE, draw.plot = TRUE)


ds.chisqr(rvar="D$sex", cvar="D$has_asthma", stvar="D$lsoa11cd",
          exclude=NULL,	useNA ="no", suppress.chisq.warnings=FALSE,
          datasources=NULL, names_region = NULL,
          force.nfilter=NULL, split = TRUE, draw.plot = FALSE)

ds.chisqr(rvar="D$sex", cvar="D$has_asthma", stvar="D$lsoa11cd",
          exclude=NULL,	useNA ="no", suppress.chisq.warnings=FALSE,
          datasources=NULL, names_region = NULL,
          force.nfilter=NULL, split = TRUE, draw.plot = TRUE)

ds.chisqr(rvar="D$sex", cvar="D$has_asthma", stvar="D$lsoa11cd",
          exclude=NULL,	useNA ="no", suppress.chisq.warnings=FALSE,
          datasources=NULL, names_region = "Liverpool",
          force.nfilter=NULL, split = TRUE, draw.plot = FALSE)

```

age/continous variable can’t be tested here (develop t-test?)

## 2. Effect of distance to greenspace

#### Is greenspace distance associated with asthma?
This is a function from dsbase.
```{r, eval=FALSE}
ds.glm(formula="has_asthma ~ greenspace_distance",family="binomial",datasources=connections
)
```


#### Does the association remain after accounting for individual characteristics?
This is a function from dsbase.
```{r, eval=FALSE}
ds.glm(formula=paste("has_asthma ~ greenspace_distance +","age + sex + ethnicity + imd_decile"
  ),family="binomial",datasources=connections
)
```


### 3. Accounting for geographical clustering

#### Question 3.1 — Is greenspace associated with asthma after accounting for LSOA clustering?
This is a function from dsbase which is based on glmer: a function from package lme4.
```{r, eval=FALSE}
ds.glmerSLMA(formula="has_asthma ~ greenspace_distance + (1|lsoa11cd)",family="binomial",datasources=connections
)
```


