---
title: "SummaryPlot"
output: github_document
date: "2026-07-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE)

```

## Function summary

Plots heatmap for lsoa level aggregate of selected variables on map. Uses modified ds.meanSdGp and MeanSdGpDS functions for lsoa level aggrgation. Default map setting from boundr package, future version accepts shapefile readable with st_read. Plotting done with library sf and ggplot2.

Current testing done on synthetic health outcome data created by RVD. 

```{r, include=FALSE}
 message("Running chunk:1")
# retrieve LSOA boundaries for Cheshire and Merseyside
# to reduce coverage, remove names from the `within_names`

cm_lsoa_sf <- boundr::bounds(
  "lsoa",
  within_level = "lad",
  within_names = c(
    "Cheshire East",
    "Cheshire West and Chester",
    "Halton",
    "Knowsley",
    "Liverpool",
    "Sefton",
    "St. Helens",
    "Warrington",
    "Wirral"
  ),
  lookup_year = 2011, # can be either 2011 or 2021
  opts = boundr::boundr_options(resolution = "BFC")
)

liverpool_sf <- cm_lsoa_sf |>
  dplyr::filter(lad22nm %in% "Liverpool")

uprn_gs_sf <- "https://pldr.org/download/2k6r3/n12/UPRN_2_1_greenspace_distances_with_coords.csv" |>
  readr::read_csv() |>
  # convert to spatial object
  sf::st_as_sf(coords = c("longitude", "latitude"), crs = 4326) |>
  # add LSOA details
  sf::st_join(cm_lsoa_sf)



qof_asthma_tbl <- "https://pldr.org/download/e6nzv/ng1/QOF_4_03_Asthma_LSOA.csv" |>
  readr::read_csv() |>
  # filter out CM LSOAs
  dplyr::filter(lsoa11 %in% cm_lsoa_sf$lsoa11cd) |>
  # filter out last year
  dplyr::filter(year == 2024)



# set seed for reproducibility
set.seed(20250603)
qof_asthma_cohort_tbl <- qof_asthma_tbl |>
  purrr::pmap(function(lsoa11, den, num, ...) {
    # recalculate pop to be a multiple of 2
    pop <- ifelse(den %% 2 != 0, den + 1, den)
    # extract UPRNs for this LSOA
    uprn_tbl <- uprn_gs_sf |>
      sf::st_drop_geometry() |>
      dplyr::filter(lsoa11cd == lsoa11) |>
      dplyr::mutate(
        dist_gs_dec = dplyr::ntile(distance_local_greenspace, 10)
      ) |>
      dplyr::arrange(dist_gs_dec)
    
    # create cohort table
    cohort_tbl <- tibble::tibble(
      lsoa11cd = lsoa11,
      sex = rep(c(0, 1), each = pop / 2),
      has_asthma = FALSE,
      uprn = sample(uprn_tbl$UPRN, size = pop, replace = TRUE)
    )
    
    # randomly assigned the outcome to `num` patients
    idx <- sample(seq_len(pop), ceiling(num), replace = FALSE)
    cohort_tbl$has_asthma[idx] <- TRUE
    
    # randomly re-assigned UPRNs with higher distances to patients with outcome
    ## create vector of probabilities, so further distances are favoured
    p <- c(0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.1, 0.2, 0.2, 0.2)
    idx_uprn_tile <- sample(1:10, ceiling(num), replace = TRUE, prob = p)
    ## subset UPRNs with match tile as per `idx_uprn_tile`
    cohort_tbl$uprn[idx] <- purrr::map(idx_uprn_tile, function(x) {
      # subset based on tile decile  
      aux <- dplyr::filter(uprn_tbl, dist_gs_dec == x)
      # extract UPRN
      aux$UPRN[sample(seq_len(nrow(aux)), 1)]
    }) |>
      purrr::list_c()
    
    # return final cohort table
    return(cohort_tbl)
  }) |>
  purrr::list_c()


qof_asthma_cohort_dist_gs_sf <- qof_asthma_cohort_tbl |>
  dplyr::left_join(uprn_gs_sf, by = c("uprn" = "UPRN", "lsoa11cd")) |>
  dplyr::arrange(lsoa11cd, uprn) |>
  sf::st_as_sf()
```

## Upload data and load back with DSLite.

```{r}
message("Running chunk:2")
library(DSLite)
library(devtools)
library(dsBase)
library(dsBaseClient)
library(dsGeospatial)
library(dsGeospatialClient)

options(nfilter.tab = 3)

dat <- qof_asthma_cohort_dist_gs_sf |> 
  dplyr::select("lsoa11cd", "has_asthma","uprn","distance_local_greenspace") |> 
  sf::st_drop_geometry()

#devtools::load_all("/Users/tkariya/Documents/DataSHIELD/dsGeospatial")


dslite.server1 <- newDSLiteServer(
  tables = list(
    uprn_asthma_green1 = dat
  )
)

dslite.server2 <- newDSLiteServer(
  tables = list(
    uprn_asthma_green2 = dat
  )
)

assign("dslite.server1", dslite.server1, envir = .GlobalEnv)
assign("dslite.server2", dslite.server2, envir = .GlobalEnv)


dslite.server1$config(defaultDSConfiguration(include=c("dsBase", "dsGeospatial")))
dslite.server1$aggregateMethod("MeanSdGpDS", "MeanSdGpDS")

dslite.server2$config(defaultDSConfiguration(include=c("dsBase", "dsGeospatial")))
dslite.server2$aggregateMethod("MeanSdGpDS", "MeanSdGpDS")




builder <- DSI::newDSLoginBuilder()

builder$append(
  server = "server_1",
  url = "dslite.server1",
  table = "uprn_asthma_green1",
  driver = "DSLiteDriver"
)

builder$append(
  server = "server_2",
  url = "dslite.server2",
  table = "uprn_asthma_green2",
  driver = "DSLiteDriver"
)


logindata <- builder$build()


conns <- DSI::datashield.login(logins = logindata, assign = FALSE)

datashield.assign.table(conns, "D", logindata)

```


Set nfilter.tab 

```{r}
nfilter.tab  <- 3
```

## Modified meanSDGPDS

modification: Convert object to numeric first and then find mean and var.


## geoheatmap function

Inputs : inputs of ds.meanSdGp and shapefile and names_region (when using boundr)
Return : mean plot of selected variable. mean can be 'split' or 'combined' across studies.





## Outputs
```{r outputs}
dsGeospatialClient::ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns)
dsGeospatialClient::ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type='split')
dsGeospatialClient::ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type='split', names_region = c("Liverpool", "Sefton"))
dsGeospatialClient::ds.geoheatmapPlot(x="D$distance_local_greenspace", y = "D$lsoa11cd",
datasources = conns)
dsGeospatialClient::ds.geoheatmapPlot(x="D$has_asthma", y = "D$lsoa11cd", datasources = conns,
type = 'split', names_region = c("Liverpool", "Sefton"))
```




